---
title: "Making Occurrence Maps"
author: "Traci P. DuBose"
date: "10/11/2020"
output: pdf_document
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(dev='jpeg')
library(tidyverse); library(sf); library(maps)
```

To investigate how many taxa had observations outside of their native range, I downloaded the Anuran range datasets from IUCN. Chloe is using these maps to help subset the native ranges for the Species Distribution Models. 

```{r}
# Anuran range maps
iucn_anuran<-read_sf('/home/tracidubose/ANURA', 'ANURA')
crs.geo <- st_crs("+proj=longlat +ellps=WGS84 +datum=WGS84")

tax_reference<-read.csv("/home/tracidubose/RCS_Anuran_input/AnuranTaxRef_20200708.csv")
anuran.taxa<-unique(tax_reference$final.taxa)

# Map of the states in the Contiguous US
states <- st_as_sf(map("state", plot = FALSE, fill = TRUE))
```
Each species should have at least one map. Subspecies are plotted separately, though their RCS is not calculated separately. This was a way to check my subsetting was working correctly and that some taxa are not accidentally mixed (like from museum collections that haven't been updated). The legend refers to a column within the IUCN range maps that describes what that range pertains to.  

```{r file locations}
PATH_SHP_HUC12 <- "/home/tracidubose/rcs_results/20201019_sp_shp_files/huc12_sp/"
PATH_SHP_L48_1km <- "/home/tracidubose/rcs_results/20201019_sp_shp_files/L48_1km/"
PATH_SHP_HyrdoBasins <- "/home/tracidubose/rcs_results/20201019_sp_shp_files/NA_hydrobasins/"
PATH_SHP_NA_1km <- "/home/tracidubose/rcs_results/20201019_sp_shp_files/NA_1km/"
```
 
```{r print the maps, echo=F, warning=F, message=F}
for(y in anuran.taxa){
  sp_l48<-st_read(PATH_SHP_L48_1km, y, quiet=T) %>% mutate(grain='buffer', ds='L48 buff')
  sp_na<-st_read(PATH_SHP_NA_1km, y, quiet=T) %>% mutate(grain='buffer', ds='NA buff')
  sp_hydro<-st_read(PATH_SHP_HyrdoBasins, y, quiet = T)%>% mutate(grain='watershed', ds='HydroBAS')
  sp_huc12<-st_read(PATH_SHP_HUC12, y, quiet = T) %>% mutate(grain='watershed', ds='HUC12')
  sts<-which(st_intersects(states, 
                      sp_huc12, sparse=F), arr.ind=T)
    
   sp_plot <- ggplot()+
    geom_sf(data=states[sts[,1],])+
    #geom_sf_text(data=states[unique(sts[,1]),], aes(label=ID))+
    geom_sf(data=sp_huc12, aes(color=ds, fill=ds))  +
    geom_sf(data=sp_hydro, aes(color=ds, fill=ds))+
    geom_sf(data=sp_na, aes(color=ds, fill=ds))+
    geom_sf(data=sp_l48, aes(color=ds, fill=ds))+
    scale_fill_viridis_d(name='Dataset', option="C", aesthetics = c('color','fill'))+
      facet_wrap(~grain, ncol=1)+
    ggtitle(y)+theme_bw() +
    theme(legend.position='bottom')#+coord_sf(xlim=sp_bb[c(1,3)], ylim=sp_bb[c(2,4)])
  print(sp_plot)
}

```

